---
title: "R Notebook"
output: html_notebook
---
##################
##################
###  APPLICATION OF THIS MODULE

##1- Create a vector city_names with the names of your favorite cities.
##Use: "Mumbai", "Delhi", "Bangalore", "Hyderabad".

```{r}

```


##2-Create a vector city_population with the population of each of these cities in millions (or approximate numbers if exact data is unavailable).
#Use: 12442373, 11007835, 8436675, 6809970.
```{r}

```


#3- Create a data frame using the two vectors above and assign it the name city_info.

```{r}

```



#4- Change the column labels in the data frame to City_Name and Population, and set the row names to (Rank 1, Rank 2, Rank 3, and Rank 4), assuming rank
#corresponds to the order of cities in terms of population.

```{r}

```


#5- Extract the Population column from the city_info data frame.
```{r}

```



#6- Calculate the total population of all cities combined.
```{r}

```



#7- Verify the structure of the data frame using a suitable function and request a summary of the data.

```{r}

```

