---
title: "R's Kickstart Workshop"
author: "Ismail Ahouari&Mehdi Moualim"
date: "2024-11-14"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r}
#install.packages("readr")
library(readr)
#install.packages("dplyr")
library(dplyr)

```

```
Let's install the required packages




```{r}
# Import dataset from the URL
url<- "https://raw.githubusercontent.com/SameulAH/R-s-Kickstart/refs/heads/main/data/1914_dataset.csv"
workshop <- read.csv(url)

workshop

```


# Display the first 5-rows of the dataset
```{r}

```
# Display the last 5-rows of the dataset

```{r}
 

```
 
 
# Displays the structure of the dataframe, including column names, types, and first few entries.
```{r}

```

# Opens the dataframe in a spreadsheet-like viewer(only works in RStudio ) using view()function 
```{r}
 
```


# Provides a summary (e.g., mean, min, max for numeric columns)
```{r}

```


Now that we have confirmed these two columns are not relevant to our dataset, drop them using the select() function and assign the result to a new variable
```{r}



```




Before going any further we need to make this column names compatible with R IDE "lieu de naissance et date ", "mère-naissance", and "conjoint-naissance" .....
we need to rename them.ex. ("lieu de naissance et date","état civil",..)


```{r}


```

```{r}
# Replace spaces and hyphens with underscores in column names
colnames(workshop) <- gsub("[.-]", "_", colnames(workshop))

# View the updated column names
print(colnames(workshop))

```


# use filter() to keep only rows where Etat_civil is "verheiratet", to keep only married individuals
```{r}

```


After renaming the columns, we can investigate the NaN values and determine the most suitable technique to handle them.
Check for NaN values in the entire dataset using is.nan()
```{r}


```


- Have a look inside the these NA values columns for any clues or pattern how to managed them by using Select( ) function   
```{r}

```





Let check What's inside these Nan-attribute in our dataset and see what are the option to deal with these imperfection in our dataset:

- If data is categorical:
Fill missing values with "unknown" or the most frequent category.
- If data is numerical:
Use mean, median, or predictive imputation based on distribution.
- If missing data is substantial:
Consider dropping the feature or using advanced imputation techniques like KNN or multiple imputation.




from view funtion we can notice the pattern that these NA values follow we can only replace the NA value with ,   and the only option left in this case is to drop them or replace NA with "unknown"

```{r}



```


Let's check if there are still any NA-value
```{r}

```

use view() function to 
```{r}

```



###################################

##Data Visualization:




#Data Visualization:

# R Data Visualization Workshop
## Exploring the 1914 Immigration Dataset

###  Basic Data Exploration

Now Lets begin the fun part of this workshop:


`
#### Exercise: Nationality Analysis
 Which are the most common nationalities in our dataset?

You can use a Bar Chart

We need to see first the unique values in that column

```{r}
unique(workshop$nationalité)

```
and we should do some big cleaning 


```{r}
# Convert to lowercase for consistency
workshop$nationalité <- tolower(workshop$nationalité)

# Remove text in parentheses (e.g., "(naturalisiert)")
workshop$nationalité <- gsub("\\s*\\(.*\\)", "", workshop$nationalité)


# Map regions to countries

# Germany
workshop$nationalité[workshop$nationalité %in% c("deutscher", "deutsche", "preusse", "preussen", 
                                                 "hessen nassauer", "sachse", "böhmer", "böhme", 
                                                 "prussian", "bayer", "german", "saxon")] <- "germany"

# France
workshop$nationalité[workshop$nationalité %in% c("franzose", "französin", "franzosin", 
                                                 "elsasser", "elsässer", "elsässerin", 
                                                 "lothringer", "lothringerin", "lothringen","french", "lorrainian", "alsatian")] <- "france"

# Italy
workshop$nationalité[workshop$nationalité %in% c("italiener", "italienerin", "tyroler", "italian")] <- "italy"

# Austria
workshop$nationalité[workshop$nationalité %in% c("oesterreicher", "oestreicher", "oestreicherin", 
                                                 "oesterreicherin", "oesterreiecher", "tyrolean","galizier", "austrian")] <- "austria"

# Belgium
workshop$nationalité[workshop$nationalité %in% c("belgier", "belgierin", "begier")] <- "belgium"

# Switzerland
workshop$nationalité[workshop$nationalité %in% c("schweitzer")] <- "switzerland"

# Netherlands
workshop$nationalité[workshop$nationalité %in% c("holländer", "dutch")] <- "netherlands"

# Luxembourg
workshop$nationalité[workshop$nationalité %in% c("luxemburgerin")] <- "luxembourg"

# Poland
workshop$nationalité[workshop$nationalité %in% c("pole", "galacian", "polish")] <- "poland"

# Russia
workshop$nationalité[workshop$nationalité %in% c("russe")] <- "russia"




# Confirm cleaned values
unique(workshop$nationalité)


# Trim whitespace
workshop$nationalité <- trimws(workshop$nationalité)

# Check the unique values again
unique(workshop$nationalité)

```




```{r}
# Simple bar plot of nationalities

   

```



 Change the orientation to horizontal bars




### Can you give us the proportion of Nationalities

Think about a pie chart
```{r}
# Count Nationalities


ggplot(workshop, aes(x = "", y = nationalité, fill = nationalité)) + geom_bar(stat = "identity", width = 1) + coord_polar(theta = "x") 
theme_void() + # Remove background grid and axis 
labs(title = "Pie Chart")
```



```



Data Cleaning état_civil
```{r}
# converting to lowercase for consistency
workshop$état_civil <- tolower(workshop$état_civil)

# Standardize marital statuses
workshop$état_civil <- ifelse(grepl("verheiratet", workshop$état_civil), "married", 
                        ifelse(grepl("ledig", workshop$état_civil), "single", 
                        ifelse(grepl("witwe|witwer", workshop$état_civil), "widowed", 
                        ifelse(grepl("geschieden", workshop$état_civil), "divorced", 
                        ifelse(grepl("wilder ehe", workshop$état_civil), "cohabiting", 
                        "other")))))

# Verify the unique values
unique(workshop$état_civil)

```

```{r}
# Stacked Bar Chart for Marital Status by Nationality



```




```








```{r}
library(ggplot2)

# Scatter plot with jitter for categorical data
library(dplyr)

# Summarize data by nationality and marital status
summary_data <- workshop %>%
  group_by(nationalité, état_civil) %>%
  summarise(count = n(), .groups = "drop")

# Bubble chart
ggplot(summary_data, aes(x = nationalité, y = état_civil, size = count)) +
  geom_point(alpha = 0.7, color = "purple") +
  scale_size_continuous(range = c(3, 10)) +
  labs(
    title = "Bubble Chart: Nationality vs Marital Status",
    x = "Nationality",
    y = "Marital Status",
    size = "Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



```







Can You please filter with just people that are german and show us a pie chart about their civil status

```{r}
# Example for one nationality (repeat for others if needed)
specific_nationality <- workshop %>% filter(nationalité == "germany")

ggplot(specific_nationality, aes(x = "", fill = état_civil)) +
  geom_bar(width = 1) +
  coord_polar(theta = "y") +
  labs(
    title = "Pie Chart: Marital Status Distribution (Germany)",
    fill = "Marital Status"
  ) +
  theme_minimal()

```





# Diamond Dataset Analysis

## Data Overview
```{r}
library(ggplot2)
data(diamonds)
```
lets see the first 6 rows of the dataset 

```{r}
head(diamonds)
```


1.How many variables are in the diamonds dataset? 
2.What are the types of each variable? 
3.How many observations are there? 

##1.Histogram:Price Distribution
```{r}

```


1. What is the most common price range for diamonds?
2. Are there any outliers in the price distribution?

##2. Box Plot : Price  by Cut
```{r}

```

1. Which cut category has the highest median price?
2. Which cut shows the most price variation?
3. Are there many outliers? What might these represent?

we can also use the simple version of ggplots and try with violin plot 
```{r}


```


##3 Scatter Plot: Price vs Carat

```{r}

```


1. What type of relationship exists between price and carat?
2. How does cut quality affect this relationship?
3. What pattern do you notice about price variation as carat size increases?

##  4. Bar Chart : Count by Clarity
```{r}

```

1. Which clarity grade is most common?
2. Which is least common?
3. Is there a pattern in the distribution?

##5. Heat Map: Price by Cut and Clarity

```{r}

```


1. Which combination of cut and clarity yields the highest average price?
2. What patterns do you notice in the heat map?
3. How does clarity affect price within each cut category?

## 6. Density Plot: Carat by Cut

```{r}



```



1. Which cut tends to have larger carat sizes?
2. How do the distributions differ between cuts?
3. Are there any unusual patterns?

## 7. Bubble Plot: Price, Carat, and Cut

```{r}

```


1. How does depth relate to price and carat?
2. Which cut shows the most variation in depth?
3. What patterns do you notice about larger diamonds?


```{r}

```

# ggplot2 Visualization Guide - Quick Reference


#==================================
# BASIC SETUP
#==================================
library(ggplot2)
library(dplyr)    # For data manipulation
library(tidyr)    # For data tidying

#==================================
#  BASIC PLOTS
#==================================

# Scatter Plot
ggplot(data, aes(x = x_var, y = y_var)) +
  geom_point()

# Line Plot
ggplot(data, aes(x = x_var, y = y_var)) +
  geom_line()

# Bar Plot
ggplot(data, aes(x = category)) +
  geom_bar()

# Histogram
ggplot(data, aes(x = numeric_var)) +
  geom_histogram(bins = 30)

#==================================
#  PLOTS
#==================================

# Scatter Plot with Trend Line
ggplot(data, aes(x = x_var, y = y_var)) +
  geom_point() +
  geom_smooth(method = "lm")

# Box Plot with Points
ggplot(data, aes(x = category, y = value)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, alpha = 0.5)

# Density Plot
ggplot(data, aes(x = numeric_var)) +
  geom_density(fill = "blue", alpha = 0.3)

#==================================
#  GROUPED VISUALIZATIONS
#==================================

# Grouped Bar Plot
ggplot(data, aes(x = category, fill = group)) +
  geom_bar(position = "dodge")

# Stacked Bar Plot
ggplot(data, aes(x = category, fill = group)) +
  geom_bar(position = "stack")

# Grouped Scatter Plot
ggplot(data, aes(x = x_var, y = y_var, color = group)) +
  geom_point()



#==================================
# CUSTOMIZATION
#==================================

# Labels and Title
ggplot(data, aes(x = x_var, y = y_var)) +
  geom_point() +
  labs(title = "Main Title",
       subtitle = "Subtitle",
       x = "X Label",
       y = "Y Label")

# Themes
ggplot(data, aes(x = x_var, y = y_var)) +
  geom_point() +
  theme_minimal()

# Custom Colors
ggplot(data, aes(x = category, fill = group)) +
  geom_bar() +
  scale_fill_brewer(palette = "Set3")

#==================================
#  PLOTS
#==================================

# Heat Map
ggplot(data, aes(x = x_var, y = y_var, fill = value)) +
  geom_tile()

# Bubble Plot
ggplot(data, aes(x = x_var, y = y_var, size = size_var)) +
  geom_point(alpha = 0.6)

# Violin Plot
ggplot(data, aes(x = category, y = value)) +
  geom_violin()

#==================================
#  COMBINING ELEMENTS
#==================================

# Box Plot with Mean Points
ggplot(data, aes(x = category, y = value)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", color = "red")

# Scatter Plot with Marginal Distributions
# Requires ggExtra package
library(ggExtra)
p <- ggplot(data, aes(x = x_var, y = y_var)) +
  geom_point()
ggMarginal(p, type = "histogram")

#==================================
#  USEFUL TIPS
#==================================

# Rotate Axis Labels
theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Add Transparency
geom_point(alpha = 0.5)

# Remove Legend
theme(legend.position = "none")

# Save Plot
ggsave("plot.png", width = 10, height = 6)

#==================================
#  COMMON DATA TRANSFORMATIONS
#==================================

# Calculate Means
data_summary <- data %>%
  group_by(category) %>%
  summarise(mean_value = mean(value))

# Create Percentage Plot
ggplot(data, aes(x = category, fill = group)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent)









