---
title: "Examples/w/Hands-on"
author: "Ismail Ahouari"
date: "2024-11-28"
output: pdf_document
---



############################
#############################
###Variables
```{r}
a <- 4
a
```



#Verify the following instructions
```{r}
a*5
(a+10)/2
a <- a+1
```


###Data types.

##my_numeric equals 42 .
#anwser here
```{r}

```


##my_character is now "universe"
#anwser here
```{r}

```

##my_logical becomes TRUE
#anwser here
```{r}

```




## Try it
```{r}
my_numeric <- 42
my_character <- "universe"
my_logical <- TRUE
the_date <- as.Date("05/29/2018", "%m/%d/%Y")
```


###Verify the data type of a variable with the
```{r}
class(my_numeric)
class(my_date)
```




###################################
##Everything is an object

```{r}
ls()
```


###Use rm() to remove an object from R's memory
## remove a single object

```{r}
rm(a)
```

### remove multiple objects

```{r}
rm(my_character, my_logical)
```


### remove a list of objects
```{r}
rm(list = c('my_date', 'my_numeric'))
```


### remove all objects
```{r}
rm(list = ls())
```


##################
##################
#### Vectors



```{r}
assign("my_vector",c(0, 3:5, 20, 0))
my_vector
```



```{r}
my_vector_2 <- c(0, 3:5, 20, 0)
my_vector_2
```
```{r}
my_vector_2[2]
length(my_vector_2)
```


```{r}
my_vectors_3 <- seq(1,10,by=2)
my_vectors_3
length(my_vectors_3)
```


```{r}
my_family <- c("Katrien", "Jan", "Leen")
my_family
```


### What happens when you try to store elements of different data types in the same vector?
```{r}

my_vector <- c("Katrien Antonio", 2, 1, 3)
names(my_vector) <- c("Name", "Profession", "Number1", "Number2")
my_vector
my_vector[1]
```



### names() function

```{r}
my_vector <- c("Katrien Antonio", "teacher")
names(my_vector) <- c("Name", "Profession")
my_vector

```




##################
##################
### Matrices


```{r}
my_matrix <- matrix(1:12, 3, 4, byrow = TRUE)
my_matrix
```


##Naming Matrix dimensions

```{r}
x <- c(R1, R2, R3)
y <- c(X,Y,Z,T)
my_matrix <- matrix(1:12, 3, 4, byrow = TRUE, dimnames=list(x,y))
my_matrix

```

## indexing in Matrix
```{r}
my_matrix[1, 3]

```

##Extract the entire 1st row

```{r}
my_matrix[1, ]
```

##Extract the entire 2nd column
```{r}

my_matrix[ ,2 ]

```

##################
##################
###  Lists

## creating a list

```{r}
my_list <- list(one = 1, two = c(1, 2), five = seq(1, 4, length=5),six = c("Katrien", "Jan"))
names(my_list)

str(my_list)
```



##################
##################
###  Dataframes

##Let’s create a dataframe
```{r}
data.frame(LETTERS [1:4],c(1,2,3,4))

```

##Naming Dataframe Dimensions
```{r}
data.frame(Alphabets=LETTERS[1:4],Numbers=c(1,2,3,4))
```


## OR we name using colnames() and rownames() function

```{r}
y_df <- data.frame(LETTERS[1:4],c(1,2,3,4))
colnames(my_df) <- c("Letters", "Digits")
rownames(my_df) <- c("Row1", "Row2", "Row3", "Row4")
```




#################################################
#################################################
#### Data Wrangling
###  select() function

```{r}
library(ggplot2)
```

```{r}
data("diamonds")
```



```{r}
selected_diamonds <- select(diamonds, color, price, carat)
selected_diamonds
```

```{r}
selected_diamonds <- diamonds %>% select(color, price, carat)
selected_diamonds
```


###  filter()

```{r}
filter(diamonds, cut == "Ideal")
```

# filter using pipe

```{r}
diamonds %>% filter(cut == "Ideal")
```

# multi-filter
```{r}
diamonds %>% filter(cut == "Ideal", price > 5000)
```




#####  mutate()

```{r}
mutate(diamonds, price_per_carat = price/carat)
```


```{r}
diamonds %>%
mutate(price_per_carat = price/carat)
```



###### summarise()

```{r}
diamonds %>% summarise(mean = mean(price), std_dev = sd(price))
```


##### group_by()

```{r}
diamonds %>% group_by(cut)%>%
 summarise( mean_price = mean(price), sd_price = sd(price))
```

#### Multistep operations

```{r}
diamonds %>%
mutate(price_per_carat = price/carat) %>%
filter(price_per_carat > 1500)
```

